Fabric-Elasticity Relationships in Osteogenesis Imperfecta (OI)

### 2021-05-18 09_Article/Simon_2021_Bone_DraftPhZ.tex
Add motion artefact grading and BMD histograms

	In conclusion, despite the reduced regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis. Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections or parts.
\end{abstract}

@@

Osteogenesis imperfecta (OI) is an inherited form of bone fragility \cite{Tournis2018}. It was estimated to concern about 1/13'500 births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}. Therefore, OI is considered as a rare metabolic bone disorder. Sillence classification \cite{Sillence1979} allows to define four main types of OI based on the clinical phenotype:
\begin{itemize}
	\item Type I: less severe
	\item Type II: lethal at most shortly after birth
	\item Type III: most severe surviving form
	\item Type IV: intermediate severity
\end{itemize}
Also known as "brittle bone disease", OI is characterized by impaired synthesis of type I collagen, leading to such brittle and fragile bones \cite{LIM2017}.\\

@@

\subsection{Image analysis}
The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Then, mask segmented images were used for further analysis.\\

@@

\subsection{Morphological Analysis}
The results of morphological analysis are summarized in Table \ref{Table2}. In the present study, the individual matching allowed to have similar group distribution with 17 females and 11 males in each group. The mean age of matched healthy is 41 $\pm$ 14 and 41 $\pm$ 15 for the matched OI. BV/TV of healthy people is higher than BV/TV of OI group with a 95\% CI of [0.016, 0.101] and p value <0.01. Similarly, trabecular number is higher in the matched healthy group with a 95\% CI of [0.099, 0.285] and a corresponding p value <0.001. The trabecular thickness do not show significant differences between groups with a p value of 0.2. In the other hand, trabecular spacing is higher in matched OI group than in healthy individuals with a p value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$). Trabecular spacing SD present to be higher in OI patients than in matched healthy with a p value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$). SMI as well as degree of anisotropy are higher for matched OI than for healthy people with p values <0.001 and of 0.02, respectively. Finally, the log transformation of coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [−0.757, −0.333].\\

@@

In conclusions, the sample analyzed in the this study have similar morphology that data reported in the literature. We find no differences in fabric-elasticity relationships between healthy and OI trabecular bone if the ROI is homogeneous enough. \citeauthor{Indermaur2021}\cite{Indermaur2021} shown that compressive behavior of OI bone tissue is similar to healthy control. If tensile and shearing behaviors agree, fabric-strength relationships will hold too. Therefore, OI bone fragility might mostly result from the decrease in BV/TV and loss of homogeneity in its organization.

\section*{Acknowledgments}
The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the building and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII scans of OI individuals. This work was internally funded by the ARTORG Center for Biomedical Engineering Research and Mereo BioPharma.

>>>>

In conclusion, despite the reduced regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis. Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections or parts.

\end{abstract}

@@

Osteogenesis Imperfecta (OI) is an inherited form of bone fragility \cite{Tournis2018}. It was estimated to concern about 1/13'500 births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}. Therefore, OI is considered as a rare metabolic bone disorder. Sillence classification \cite{Sillence1979} allows to define four main types of OI based on the clinical phenotype:
\begin{itemize}
	\item Type I: less severe
	\item Type II: lethal at most shortly after birth
	\item Type III: most severe surviving form
	\item Type IV: intermediate severity
\end{itemize}
Also known as "brittle bone disease", OI is characterized by impaired synthesis of type I collagen, leading to such brittle and fragile bones \cite{LIM2017}.\\

@@

\subsection{Image analysis}
The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Then, mask segmented images were used for further analysis. \\

@@

\subsection{Morphological Analysis}
The results of morphological analysis are summarised in Table \ref{Table2}. In the present study, the individual matching allowed to have similar group distribution with 17 females and 11 males in each group. The mean age of matched healthy is 41 $\pm$ 14 and 41 $\pm$ 15 for the matched OI. BV/TV of healthy people is higher than BV/TV of OI group with a 95\% CI of [0.016, 0.101] and p value <0.01. Similarly, trabecular number is higher in the matched healthy group with a 95\% CI of [0.099, 0.285] and a corresponding p value <0.001. The trabecular thickness do not show significant differences between groups with a p value of 0.2. In the other hand, trabecular spacing is higher in matched OI group than in healthy individuals with a p value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$). Trabecular spacing SD present to be higher in OI patients than in matched healthy with a p value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$). SMI as well as degree of anisotropy are higher for matched OI than for healthy people with p values <0.001 and of 0.02, respectively. Finally, the log transformation of coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [−0.757, −0.333].\\

@@

In conclusions, the sample analysed in the this study have similar morphology than data reported in the literature. We find no differences in fabric-elasticity relationships between healthy and OI trabecular bone if the ROIs are homogeneous enough. \citeauthor{Indermaur2021}\cite{Indermaur2021} showed that compressive behaviour of OI bone tissue is similar to healthy control. If tensile and shearing behaviours agree, fabric-strength relationships will hold too. Therefore, OI bone fragility might mostly result from the decrease in BV/TV and loss of homogeneity in its macroscopic organisation.

\section*{Acknowledgments}
The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the buildingi and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII scans of OI individuals. This work was internally funded by the ARTORG Center for Biomedical Engineering Research and Mereo BioPharma.
